Automatic Matching of Bullet Lands

### 2016-05-23 revision/response.Rmd

---
title: "Response to reviewers"
author: "Heike Hofmann, Eric Hare, Alicia Carriquiry"
date: "April 14, 2016"
output: html_document
---

# Overall TODO

1. Read the paper start to finish
2. Look at boxplots vs densities (maybe use jittered boxplots?)
3. Talk to Alen, send paper, ask to add to intro? Add as author?

@@

***xie:2009 first sentence: "Every firearm has individual characteristics that are as unique to it as fingerprints are to human beings." the paper itself doesn't back up any of these claims. very disappointing***

***leon:2006 first sentence: "Fired bullets bear striation marks that can be thought of as a ‘‘fingerprint’’ left by the firearm." based on multiple 2d images. They extract a 1d 'signature'***

5. page 7- figure 4- What exactly are you bootstrapping? It is unclear in the paper whether you are bootstrapping residual or bullets or something else entirely. Please clarify.

***We have decided to remove the bootstrap confidence intervals for the loess fit residuals because they did not add enough information.***

@@

11. In the conclusions section the authors mention an assumption of "uniqueness". In general forensic science have moved beyond these questions- see the work of Saks and Kaye for details.

***The reviewer is absolutely correct, but we are talking about a different ``uniqueness'' concept here.  The discipline of firearm and toolmark examining relies on two foundational assumptions: that every gun leaves a unique set of markings and that the marks are reproducible. This is still the credo among tool and firearms examiners.  The reviewer is referring to the fact that in testimony, forensic scientists are beginning to move away from declaring that, for example, an individual gun, to the exclusion of all others, was the one that fired the bullet.  The work we present addresses the fundamental assumption of uniqueness, and will hopefully help move the practice away from statements of unique source.***

12. A discussion on how the authors view the use of their matching algorithm would be useful to the forensic community. For example are they intending to use this to build a closed set identification system similar to IAFIS or are them intending this to work in a manner similar to a biometric verification system? etc.

***The problem with existing closed set systems is that it is unclear how the matching is exactly done and what statistical properties it has (number of parameters, distribution of them, sensitivities). Repeatability and reproducibility is a huge issue in this area. 'Matching' is also not just a single step - it consists, as we have laid out in the paper, of a series of steps, which all deserve attention. While it was pointed out to us to use 'standard methods' - there are no actual standards in place. The status quo is NIBIN, which is not accessible to anyone outside of law enforcement. There might be a set of alternatives, and some evidence as to which methods might work better in some situations than others, but there is a lack of an in-depth discussion of the alternatives at each step as well as an overall discussion. In this manuscript we are detailing one route to divvy up the process and provide the code for others to reproduce our findings.***

***TODO: Add some of this discussion to the intro + conclusion***

@@

14. The authors would need to study (or at least discuss extensively) the ability of their method to capture the variability in bullet profiles over time (at interval of several dozen of bullets fired by the same weapon, when the weapon has been cleaned, or when the bullet was degraded by hitting a target). In addition, a dataset of 10 firearms (even consecutively manufactured) does not provide any information on the rarity of a particular bullet profile, and therefore does not address the second probability statement. Alternatively, the proposed metrics could be tested in a database searching situation, which has not been done.

***At the time of the submission of this manuscript, the 35 bullets from 10 barrels from Hamby set 252 were the only ones accessible to the public. They are part of Dr Zheng's efforts at NIST to provide a ballistics reference database. In the future more results will be available, but until then we are unable to provide a data-informed discussion of the very important issues raised by the reviewer.***

15. On a related subject, I believe that a random forest provides the probability that two bullets were fired by the same gun given the set of observations (Fig 19), not the probability of interest: the probability of two sets of observations given that the two bullets were fired by the same gun. The authors may want to revise some of the statements in their discussion and conclusion.

***The reviewer is correct in stating that two probabilities of interest are the probability of observing the data under the "same source" (same gun) hypothesis or under the "different source" hypothesis.  An empirical approach to get at those two probabilities is to derive the distribution of scores when two bullets are known to have been fired by the same gun and the distribution of scores when two bullets are known to have been fired by different guns.  We have actually derived those to distributions and include them in the revised manuscript.***

@@

16. I am concerned by the assumption of independence of between the different lands on a given bullet that the authors seems to use in their error rates calculations (and ROC curves). The assumption seems unreasonable since as soon as a land has been paired between two bullets, it seems that it becomes more likely to pair the other lands. In other words, I believe that the authors need to redo their error rate measurements, but using bullets as their experimental units and not lands. I am also concerned by the overwhelming majority of non-matches, this must certainly artificially increase the rate of correct exclusion and distort the data.

***Assuming independence between the lands of each barrel is certainly not realistic in practice, however, by assuming independence we make the task of matching harder for the algorithm. This also has the benefit of potentially being more useful for degraded data, where we may not have a full set of six groove-to-groove scans. Based on the actual numerical values of features derived from each of the lands, an independence assumption is not unrealistic, because the features do not exhibit any clustering associated to barrel. The technique suggested by the reviewer could be solved using a hierarchical approach to matching, and we agree with the assessment that it would likely decrease the error rates.***

***TODO: We will figure out whether there is clustering associated with the barrel. Add discussion of why we are comparing land to land rather than bullet to bullet***

### 2016-05-26 revision/response.Rmd

***The reviewer is correct in stating that two probabilities of interest are the probability of observing the data under the "same source" (same gun) hypothesis or under the "different source" hypothesis.  An empirical approach to get at those two probabilities is to derive the distribution of scores when two bullets are known to have been fired by the same gun and the distribution of scores when two bullets are known to have been fired by different guns.  We have actually derived those to distributions and include them in the revised manuscript.***

16. I am concerned by the assumption of independence of between the different lands on a given bullet that the authors seems to use in their error rates calculations (and ROC curves). The assumption seems unreasonable since as soon as a land has been paired between two bullets, it seems that it becomes more likely to pair the other lands. In other words, I believe that the authors need to redo their error rate measurements, but using bullets as their experimental units and not lands. I am also concerned by the overwhelming majority of non-matches, this must certainly artificially increase the rate of correct exclusion and distort the data.

***Assuming independence between the lands of each barrel is certainly not realistic in practice, however, by assuming independence we make the task of matching harder for the algorithm. This also has the benefit of potentially being more useful for degraded data, where we may not have a full set of six groove-to-groove scans. Based on the actual numerical values of features derived from each of the lands, an independence assumption is not unrealistic, because the features do not exhibit any clustering associated to barrel. The technique suggested by the reviewer could be solved using a hierarchical approach to matching, and we agree with the assessment that it would likely decrease the error rates.***

***TODO: We will figure out whether there is clustering associated with the barrel.***

### 2016-05-26 imaging-paper.Rnw
Continue working on many of the changes

The purpose of our paper is to present an automatic matching routine that allows for a completely objective assessment of the quality of a match. 
The original paper on the complete Hamby study already reports the successful use of several computer-assisted methods; aside from a zero false positive rate, no further details on a false-negative error rate for bullets are given nor are error rates for land-to-land matches mentioned. 
% Hamby study: automated systems were successfully used for matching:
%  Intelligent Automation's SciClops - Dr. Ben Bachrach (Maryland, United States);
% Automated Land Identification System (ALIS) - Mr. Tsuneo Uchiyama (Tokyo, Japan);
% Integrated Ballistics Identification System (IBIS) – Mr. Robert Thompson (California, United States);
% BulletTRAX-3DTM - Forensic Technology Scientists (Montreal, Canada) (2 sets)

@@

We have presented an algorithm which detects the most prominent but least relevant structure of a bullet from a firearms identification perspective, removes these features, and produces residuals which allow for the easy identification of markings. We have generalized this algorithm to align the residuals from two bullets to automatically determine whether they are matches. We created a random forest model using many of the features to provide a probabilistic assessment of the quality of a match, along with the most relevant features. %Finally, we have provided an easy-to-use open-source backend, and web frontend to many of these features.

>>>>

The purpose of our paper is to present an automatic matching routine that allows for a completely objective assessment of the quality of a match. \eh{We have chosen to perform the matching on a land-to-land level, rather than bullet-to-bullet. Although doing so introduces an implicit assumption of independence between lands, assuming independence only serves to make the task more challenging. XXX need more?}
The original paper on the complete Hamby study already reports the successful use of several computer-assisted methods; aside from a zero false positive rate, no further details on a false-negative error rate for bullets are given nor are error rates for land-to-land matches mentioned. 
% Hamby study: automated systems were successfully used for matching:
%  Intelligent Automation's SciClops - Dr. Ben Bachrach (Maryland, United States);
% Automated Land Identification System (ALIS) - Mr. Tsuneo Uchiyama (Tokyo, Japan);
% Integrated Ballistics Identification System (IBIS) – Mr. Robert Thompson (California, United States);
% BulletTRAX-3DTM - Forensic Technology Scientists (Montreal, Canada) (2 sets)

@@

We have presented an algorithm which detects the most prominent but least relevant structure of a bullet from a firearms identification perspective, removes these features, and produces residuals which allow for the easy identification of markings. We have generalized this algorithm to align the residuals from two bullets to automatically determine whether they are matches. We created a random forest model using many of the features to provide a probabilistic assessment of the quality of a match, along with the most relevant features. \eh{Matching bullets is clearly not a single step, but rather a composition of a number of data analysis tasks which each deserve attention. With no scientific standards in place, our intent is to explain one approach to addressing these tasks, while documenting all steps and providing all code so other researchers and forensic scientists can reproduce and expand on our findings.}
